---
title: 'Two independent samples t-test'
author: Danilo Sarti \newline \texttt{danilo.sarti@mu.ie}   \newline \vspace{1cm}
  \newline \vspace{1cm}
  https://danilosarti.github.io/
  \newline PRESS RECORD 
output:
  beamer_presentation:
    includes:
      in_header: header.tex
classoption: "aspectratio=169"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(dev = 'pdf', fig.height =4)
options(width = 40)
par(mar=c(3,3,2,1), mgp=c(2,.7,0), tck=-.01,las=1)
```

## Learning outcomes

- Know the motivation and when to use the t-test
- Explore a dataset to answer a question
- Know how to apply the test with R
- Extract conclusions from the test

## Motivation 


- Suppose we work for a brand that is producing a new drug for weight loss
- For some reason, you are asked to address the question: Do men and women differ in terms of weight on average? This will become the hypothesis to be tested (check previous classes!!)
- A statistical test can answer this question: the t-test (developed by Student) 
- The t-test allows us to compare the means of two independent groups
- Interesting fact: The real Student was William Gosset (Guinness brewer!!)


## Loading packages to help

- As we learned in the last lab sessions, let's load some R packages that will help us!! All code: https://github.com/danilosarti/t_test_class

```{r, warning=FALSE, message=FALSE}
library(tidyverse) #to help tidy data
library(ggpubr) #to produce some good graphs
library(rstatix) #to perform statistical tests in a easy manner!!!
```

## The data. 

- Suppose that your boss also provided you with a data set containing the weights of 20 men and 20 women. 
- That data is stored in the R package datarium we used in previous classes. Let's take a look at the data.

```{r}
# Load the data which is actually stored in the datarium R package.
data("genderweight", package = "datarium")
# Show a sample of the data by group
genderweight %>% sample_n_by(group, size = 2)
```

## Data Visualisation

- As we learned is always helpful to visualize the data!!!
- Let us produce some boxplots with the code!!!  

```{r}
df=genderweight %>%
  group_by(group)
p=ggboxplot(df, x = "group", y = "weight",
                color = "group", palette =c("#00AFBB", "#E7B800"),
                add = "jitter", shape = "group")
```

## What do we see in terms of averages?
```{r}
p ## printing our nice plot!!
```


## Data summary

- We can easily calculate some summary statistics 
- Is there evidence that the means are different?
- How can we really test that? Solution: usage of a statistical test!!!

```{r}
genderweight %>%
  group_by(group) %>%
  get_summary_stats(weight, type = "mean_sd")
```

## Recalling tests

- All tests have assumptions
- They are associated with something we call statistics (check previous material!!)
- They return a p-value associated with the hypothesis tested ($H_0$). If the p-value<0.05, or another value we may consider proper, it does not support the hypothesis
- They may have some additional features called parameters! (check the previous classes!!)

## t-test and t-test

- We have, indeed two forms for the t-test
- The original Student test assumes that the variance (variability-sd^2) of the two groups are equal. Not so real for our data set!!
- The Welch version: Does not consider variances to be identical. It fits our case more!!
- Other names: independent t-test,independent samples t-test, unpaired t-test or unrelated t-test.
- The test considers the probability distribution of the differences between means of women and men

## Proceedig the t-test in a pipe-friendly manner (tidyverse)

- Welch version of the test!! $H_o=$are the average weight of men and women different? 
- We are using commands from the package rstatix
- We see The results above show the following components:

    
    .y.: the y variable used in the test -- weight.
    group1,group2: the compared groups in the tests.
    statistic: Test statistic used to compute the p-value. 
    df: degrees of freedom. A parameter associated to the test
    p: p-value. Is it less than 0.05?



```{r}
stat.test <- genderweight %>% 
  t_test(weight ~ group) %>%
  add_significance()
data.frame(stat.test) # to show complete results!
```

## Student t-test 

- If you want to proceed with the original t-student test, use: `var.equal = TRUE`
- Do the final result regarding the means remains the same? $H_o=$are the average weight of men and women different? 

```{r}
stat.test2 <- genderweight %>%
  t_test(weight ~ group, var.equal = TRUE) %>%
  add_significance()
data.frame(stat.test) # to show complete results!
```

## How about the confidence interval for difference of the means?

- If you want to obtain also de confidence interval for difference of the means you can use the basic r function t.test!! 

```{r}
# Compute t-test
res <- t.test(weight ~ group, data = genderweight)
```

## 
- conf.int is the confidence interval of the means difference at 95% (conf.int = [-24.5314, -20.1235])
```{r}
res
```



## We need to answer our boss!!
- Remember, your boss does not know data science. We need to inform him anyways!!
- The mean of the women is 63.5 with an associated standard deviation (sd) of 2.03; for men, the mean is 85.8(sd=4.3). A statistical test (Welch's t-test) returns a p-value <0.0001, which suggests that the weight of men and women differ on average. The results can also be visualized as follows: 

## 
```{r, output=FALSE}
# Create a black and whit box-plot -- useful for some publications!!
bxp <- ggboxplot(
  genderweight, x = "group", y = "weight", 
  ylab = "Weight", xlab = "Groups", add = "jitter"
  )

# Add p-value and significance levels obtained with code in previous slides!!!!
stat.test <- stat.test %>% add_xy_position(x = "group")

```

##
```{r}
bxp + 
  stat_pvalue_manual(stat.test, tip.length = 0) +
  labs(subtitle = get_test_label(stat.test, detailed = TRUE))
```

